fGFPCA dynamic prediction

### 2023-03-01 f-GFPCA.Rmd
have trouble knitting computation time

Table below include time spent only for prediction.

@@

```{r}
# table displaying predictive performance
tb_pred_eval <- tibble(
  c("195", "395", "595"),
  ise_fgfpca, ise_adglmm, auc_fgfpca, auc_adglmm, comp_t_fgfpca, comp_t_adglmm)
colnames(tb_pred_eval) <- 
  c("Maximum observed time", rep(c("fGFPCA", "GLMMadaptive"), 3))
```

```{r, results='asis'}
tb_pred_eval %>% kable(digits = 3, align = "lcccccc") %>%
  kable_styling(full_width = F) %>%
  add_header_above(c(" "=1, "ISE"=2, "AUC "=2, "Prediction time"=2))
```

@@

- Based on posterior likelihood
- Standard deviation from this method ended up being very big
- Also does not change from in-sample to out-of-sample, nor across subjects

3. Bootstrap

```{r, warning=FALSE}
source(here("Code/OutSampPredInterv.R"))

@@

```{r}
pred_std_mle <- data.frame(mid)

for(t in c(195, 395, 595)){
  mle_xi <- out_samp_dyn_pred(df_new = df %>% filter(id==1 & sind_inx<=t) %>%
                           select(-eta_i),fpca_fit = fpca_fit)$score_out
  I_mat <- -gradient(f = llh_div, x = mle_xi, 
                     df_new = df %>% filter(id==1 & sind_inx<=195) %>% select(-eta_i),
                   fpca_fit=fpca_fit)
  std <- sqrt(diag(fpca_fit$efunctions %*% solve(I_mat) %*% t(fpca_fit$efunctions))+fpca_fit$sigma2)

  pred_std_mle <- data.frame(pred_std_mle, std)
}
pred_std_mle <- pred_std_mle %>%  pivot_longer(2:4, values_to = "std_mle", names_to = "type") %>%
  mutate(type = factor(type, levels = c("std", "std.1", "std.2"),
                       labels=c("Up to 195", "Up to 395", "Up to 595")))
```

```{r}
out_pred_subj1 %>%
  left_join(pred_std_eb, by = c("mid", "type")) %>% 
  mutate(eta_upper_eb = eta+std_eb,
         eta_lower_eb = eta-std_eb) %>%
  left_join(pred_std_mle, by = c("mid", "type")) %>% 
  mutate(eta_upper_mle = eta+std_mle,
         eta_lower_mle = eta-std_mle) %>%
  ggplot()+
    geom_line(aes(x=mid, y=true), col="black")+
    geom_line(aes(x=mid, y=eta, col="Estimated"))+
    geom_line(aes(x=mid, y=eta_upper_eb, col="EB"), linetype="dashed")+
    geom_line(aes(x=mid, y=eta_lower_eb, col="EB"), linetype="dashed")+
  geom_line(aes(x=mid, y=eta_upper_mle, col="MLE"), linetype="dashed")+
    geom_line(aes(x=mid, y=eta_lower_mle, col="MLE"), linetype="dashed")+
    facet_wrap(~type)
```

>>>>

@@

```{r}
# table displaying predictive performance
tb_pred_eval <- tibble(
  c("195", "395", "595"),
  ise_fgfpca, ise_adglmm, auc_fgfpca, auc_adglmm)
colnames(tb_pred_eval) <- 
  c("Maximum observed time", rep(c("fGFPCA", "GLMMadaptive"), 3))

# print(tb_pred_eval)
# print(comp_t_fgfpca)
# print(comp_t_adglmm)
```

```{r, results='asis'}
tb_pred_eval  %>% 
  kable(format = "html", digits = 3, align = "lcccccc", 
        table.attr = "style = \"color: black;\"") %>%
  kable_styling(full_width = F) %>%
  add_header_above(c(" "=1, "ISE"=2, "AUC "=2))
```

@@

- Based on posterior likelihood
- Also ignore randomness from FPCA estimates

$$Var(\hat{\boldsymbol{\eta}}-\boldsymbol{\eta})=\boldsymbol{\Phi}I(\hat{\boldsymbol{\xi}})\boldsymbol{\Phi}^T+\hat{\sigma_{\epsilon}}^2\boldsymbol{I}_m $$
- Standard deviation would change across subjects.

```{r}
source(here("Code/OutSampPredInterv.R"))

@@

```{r}
df_std_mle_subj1 <- data.frame(mid, 
                               t195 = pred_std_mle(df %>% 
                                                     filter(id==1 & sind_inx<=195) %>%
                                                     select(-eta_i),
                                                   fpca_fit),
                               t395 =  pred_std_mle(df %>% 
                                                     filter(id==1 & sind_inx<=395) %>%
                                                     select(-eta_i),
                                                   fpca_fit),
                               t595 =  pred_std_mle(df %>% 
                                                     filter(id==1 & sind_inx<=595) %>%
                                                     select(-eta_i),
                                                   fpca_fit)) %>%
  pivot_longer(2:4, values_to = "std_mle", names_to = "type") %>%
  mutate(type = factor(type, levels = c("t195", "t395", "t595"),
                       labels=c("Up to 195", "Up to 395", "Up to 595")))

df_std_mle_subj2 <- data.frame(mid, 
                               t195 = pred_std_mle(df %>% 
                                                     filter(id==2 & sind_inx<=195) %>%
                                                     select(-eta_i),
                                                   fpca_fit),
                               t395 =  pred_std_mle(df %>% 
                                                     filter(id==2 & sind_inx<=395) %>%
                                                     select(-eta_i),
                                                   fpca_fit),
                               t595 =  pred_std_mle(df %>% 
                                                     filter(id==2 & sind_inx<=595) %>%
                                                     select(-eta_i),
                                                   fpca_fit))%>%
  pivot_longer(2:4, values_to = "std_mle", names_to = "type") %>%
  mutate(type = factor(type, levels = c("t195", "t395", "t595"),
                       labels=c("Up to 195", "Up to 395", "Up to 595")))

df_std_mle_subj3 <- data.frame(mid, 
                               t195 = pred_std_mle(df %>% 
                                                     filter(id==3 & sind_inx<=195) %>%
                                                     select(-eta_i),
                                                   fpca_fit),
                               t395 =  pred_std_mle(df %>% 
                                                     filter(id==3 & sind_inx<=395) %>%
                                                     select(-eta_i),
                                                   fpca_fit),
                               t595 =  pred_std_mle(df %>% 
                                                     filter(id==3 & sind_inx<=595) %>%
                                                     select(-eta_i),
                                                   fpca_fit))%>%
  pivot_longer(2:4, values_to = "std_mle", names_to = "type") %>%
  mutate(type = factor(type, levels = c("t195", "t395", "t595"),
                       labels=c("Up to 195", "Up to 395", "Up to 595")))

df_std_mle_subj4 <- data.frame(mid, 
                               t195 = pred_std_mle(df %>% 
                                                     filter(id==4 & sind_inx<=195) %>%
                                                     select(-eta_i),
                                                   fpca_fit),
                               t395 =  pred_std_mle(df %>% 
                                                     filter(id==4 & sind_inx<=395) %>%
                                                     select(-eta_i),
                                                   fpca_fit),
                               t595 =  pred_std_mle(df %>% 
                                                     filter(id==4 & sind_inx<=595) %>%
                                                     select(-eta_i),
                                                   fpca_fit))%>%
  pivot_longer(2:4, values_to = "std_mle", names_to = "type") %>%
  mutate(type = factor(type, levels = c("t195", "t395", "t595"),
                       labels=c("Up to 195", "Up to 395", "Up to 595")))

```{r, warning=FALSE,fig.height=3, fig.width=10}
out_pred_subj1 %>%
  left_join(pred_std_eb, by = c("mid", "type")) %>% 
  mutate(eta_upper_eb = eta+std_eb,
         eta_lower_eb = eta-std_eb) %>%
  left_join(df_std_mle_subj1, by = c("mid", "type")) %>% 
  mutate(eta_upper_mle = eta+std_mle,
         eta_lower_mle = eta-std_mle) %>%
  ggplot()+
    geom_line(aes(x=mid, y=eta, col="Estimated"))+
    geom_line(aes(x=mid, y=eta_upper_eb, col="EB"), linetype="dashed")+
    geom_line(aes(x=mid, y=eta_lower_eb, col="EB"), linetype="dashed")+
  geom_line(aes(x=mid, y=eta_upper_mle, col="MLE"), linetype="dashed")+
    geom_line(aes(x=mid, y=eta_lower_mle, col="MLE"), linetype="dashed")+
    facet_wrap(~type)

out_pred_subj2 %>%
  left_join(pred_std_eb, by = c("mid", "type")) %>% 
  mutate(eta_upper_eb = eta+std_eb,
         eta_lower_eb = eta-std_eb) %>%
  left_join(df_std_mle_subj2, by = c("mid", "type")) %>% 
  mutate(eta_upper_mle = eta+std_mle,
         eta_lower_mle = eta-std_mle) %>%
  ggplot()+
    geom_line(aes(x=mid, y=eta, col="Estimated"))+
    geom_line(aes(x=mid, y=eta_upper_eb, col="EB"), linetype="dashed")+
    geom_line(aes(x=mid, y=eta_lower_eb, col="EB"), linetype="dashed")+
  geom_line(aes(x=mid, y=eta_upper_mle, col="MLE"), linetype="dashed")+
    geom_line(aes(x=mid, y=eta_lower_mle, col="MLE"), linetype="dashed")+
    facet_wrap(~type)

out_pred_subj3 %>%
  left_join(pred_std_eb, by = c("mid", "type")) %>% 
  mutate(eta_upper_eb = eta+std_eb,
         eta_lower_eb = eta-std_eb) %>%
  left_join(df_std_mle_subj3, by = c("mid", "type")) %>% 
  mutate(eta_upper_mle = eta+std_mle,
         eta_lower_mle = eta-std_mle) %>%
  ggplot()+
    geom_line(aes(x=mid, y=eta, col="Estimated"))+
    geom_line(aes(x=mid, y=eta_upper_eb, col="EB"), linetype="dashed")+
    geom_line(aes(x=mid, y=eta_lower_eb, col="EB"), linetype="dashed")+
  geom_line(aes(x=mid, y=eta_upper_mle, col="MLE"), linetype="dashed")+
    geom_line(aes(x=mid, y=eta_lower_mle, col="MLE"), linetype="dashed")+
    facet_wrap(~type)

out_pred_subj4 %>%
  left_join(pred_std_eb, by = c("mid", "type")) %>% 
  mutate(eta_upper_eb = eta+std_eb,
         eta_lower_eb = eta-std_eb) %>%
  left_join(df_std_mle_subj4, by = c("mid", "type")) %>% 
  mutate(eta_upper_mle = eta+std_mle,
         eta_lower_mle = eta-std_mle) %>%
  ggplot()+
    geom_line(aes(x=mid, y=eta, col="Estimated"))+
    geom_line(aes(x=mid, y=eta_upper_eb, col="EB"), linetype="dashed")+
    geom_line(aes(x=mid, y=eta_lower_eb, col="EB"), linetype="dashed")+
  geom_line(aes(x=mid, y=eta_upper_mle, col="MLE"), linetype="dashed")+
    geom_line(aes(x=mid, y=eta_lower_mle, col="MLE"), linetype="dashed")+
    facet_wrap(~type)
```
